---
title: "Confidence Intervals for Variances from lm"
author: "Kevin"
date: "Friday, April 03, 2015"
output: 
  html_document:
    theme: cerulean
---

```{r knitr-setup, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      tidy = FALSE, 
                      fig.height = 5, 
                      fig.width = 8, 
                      fig.align = "center", 
                      size = "small", 
                      digits = 3, 
                      comment = "#>")
options(show.signif.stars = F, 
        digits = 3)
library(katexR)
```


## Very Quick Example

Simulated 95% confidence interval for the residual standard error from an `lm` model for the mtcars dataset.

```{r mtcars_sim}
m1 <- lm(mpg ~ wt, data = mtcars)
library(arm)
sims <- sim(m1, 1000)
quantile(sims@sigma, probs = c(.025, .975))
```

## The Linear Model

The following linear model can be fit in R with the `lm` function.

$mpg_i = \beta_{0} + \beta_{1} wt_{i} + \epsilon_{i}$

$\epsilon_i \sim N(0, \sigma^2)$

```{r mtcars_lm}
m2 <- lm(mpg ~ wt, data = mtcars)
summary(m2)
```

In this model, there are three parameters: $\beta_0$, $\beta_1$, and $\sigma$.  R estimates these parameters as 37.3, -5.3, and 3.0 respectively.  

## The `sim` function

The `arm` package was written by Andrew Gelman and Yu-Song Su to complement Gelman's *Data Analysis Using Regression and Multilevel/Hierarchical
Models* book.  In Chapter 7 of this book, they discuss how to use simulation for inferences in regression context.

To aid with these inferences, they provided the `sim` function in the package.  This function simulates estimated parameters from the fitted model.  In the code below, I simulate $\beta_0$, $\beta_1$, and $\sigma$ 1000 times.  The `str` command reveals the structure of an R object.  We can see that the simulated values are stored in two separate parts of the `sims` object.  The estimated regression coefficients ($\beta_0$ and $\beta_1$) are stored in a matrix named `coef`, and the estimated variabilities (just $\sigma$ in this example) are named `sigma`.

```{r lm_sims}
library(arm)
sims <- sim(m2, 1000)
str(sims)
```

## Confidence Intervals for $\sigma$

These simulations can be used to provide a summary of the model's suggested values of the parameters.

```{r sims_summary}
summary(sims@sigma) ## the @ sign is used because sim uses S4 class objects
```

We can visualize the distribution of simulated values.

```{r sims_density}
library(ggplot2)
qplot(sims@sigma, geom = "density", fill = I("lightblue")) + 
  labs(x = expression(sigma), 
       title = "Simulated Distribution of Residual Standard Error")
```

To generate a confidence interval for $\sigma$, we can just extract the middle 95% of this distribution using the `quantile` function.

```{r sims_ci}
quantile(sims@sigma, probs = c(.025, .975))
```





